Estimation of the perceived time of presence of sources in urban acoustic environment using deep learning techniques

### 2019-03-28 paper/paperDraft.tex
input for intro

\section{Introduction}
\label{sec:intro}
%- Du bruit au soundscape
%- La pleasantness premier axe sur les affects liés au soundscape
%- On a des modèles de pleasantness perceptifs qui font intervenir des temps de présence des sources
%- Peut-on à partir de la mesure physique faire des modèles de pleasantness sur la base d'indicateurs physiques?
%- Idée: si l'on fait de la reconnaissance de sources, on peut faire un meilleur modèle physique que les précédents
%- Première étape: On vérifie que le corpus créé est proche d'un corpus réel
%- Sur le corpus on fait de l'apprentissage profond
%- On compare le nouveau modèle physique deep en comparaison aux modèles physiques old school Cartazur (dimension purement niveau sonore) et Grafic (niveau sonore + Fréquence/sources)
In urban areas, the decreasing quality of sound environments resulting from urbanization is a major cause of annoyance, and has been identified as an important health factor. Consequently, the 2002/49/CE European directive requires the development of noise maps in large cities. These maps aim at providing useful tools for urban planning and noise reduction purposes. As an alternative to the noise-only paradigm used until recently, the soundscape approach defined in the ISO 12913-1 (CITE) norm as "the acoustic environment as perceived and understood and/or experienced by people and/or society, in context" extends the characterization of sound environments by considering the positive influence of some of its components.\\

@@

The perceptual characterization of soundscapes requires subjective inputs, and thus cannot directly be applied to the construction of perceptually motivated noise maps. With the advent of the Internet of Things and the availability of low-cost sensor networks (CITE), predicting pleasantness from acoustic measurements is an interesting and ongoing area of research as an alternative to systematic human quality assessment. Ricciardi et al. (CITE) present a physical model of pleasantness using only variations of the measured sound level. Though, because these indicators are not representative of the composition of the sound scene, the model underperforms in comparison to predictions using perceptual assessments. Aumond et al. (CITE) explore new indicators based on spectral variations in recordings of sound scenes. These indicators are designed to represent the activity of specific sound sources, specifically human voices and birds, but their discriminative properties are limited. Following the recent advances in machine learning, it may be possible to better approximate the perception of sound sources contributing to pleasantness by developing new indicators relying on source recognition models. However, large amounts of annotated data are needed for these architectures to learn to extract relevant information from recordings. To our knowledge such databases do not currently exist in the literature. A possible alternative is the use of simulated corpora (CITE) where sound scenes are composed from semi-randomly generated annotations.\\

The aim of this research is two-fold. First, a perceptual test is conducted where simulated sound scenes are compared to recordings to validate their use in the context of pleasantness prediction. Next, a deep learning architecture is designed and trained that performs multi-label sound source recognition. Estimations of pleasantness using these predictions are compared to existing models from physical indicators.

>>>>

\section{Introduction}
\label{sec:intro}
%- Du bruit au soundscape
%- La pleasantness premier axe sur les affects liés au soundscape
%- On a des modèles de pleasantness perceptifs qui font intervenir des temps de présence des sources
%- Peut-on à partir de la mesure physique faire des modèles de pleasantness sur la base d'indicateurs physiques?
%- Idée: si l'on fait de la reconnaissance de sources, on peut faire un meilleur modèle physique que les précédents
%- Première étape: On vérifie que le corpus créé est proche d'un corpus réel
%- Sur le corpus on fait de l'apprentissage profond
%- On compare le nouveau modèle physique deep en comparaison aux modèles physiques old school Cartazur (dimension purement niveau sonore) et Grafic (niveau sonore + Fréquence/sources)
In urban areas, the decrease of quality of sound environments resulting from urbanization is a major cause of annoyance, and has been identified as an important health factor. Consequently, the 2002/49/CE European directive requires the development of noise maps in large cities. These maps aim at providing useful tools for urban planning and noise reduction purposes. As an alternative to the noise-only paradigm used until recently, the soundscape approach defined in the ISO 12913-1 (CITE) norm as "the acoustic environment as perceived and understood and/or experienced by people and/or society, in context" extends the characterization of sound environments by considering the positive influence of some of its components.\\

@@

Though, the perceptual characterization of soundscapes requires subjective inputs, and thus cannot directly be applied to the production of perceptually motivated noise maps. With the advent of the Internet of Things and the availability of low-cost sensor networks (CITE), it may be feasible to predict pleasantness or other perceptual factors from acoustic measurements.

Bridging the gap between raw data issued from the sensors to reasonable estimates of perceptual attributes is an interesting and ongoing area of research. Ricciardi et al. (CITE) present a physical model of pleasantness using only variations of the measured sound level. Though, because these indicators are not representative of the composition of the sound scene, the model underperforms in comparison to predictions using perceptual assessments. Aumond et al. (CITE) explore new indicators based on spectral variations in recordings of sound scenes. These indicators are designed to model the activity of specific sound sources, specifically human voices and birds, but their discriminative properties are limited.

In order to tackle this issue, we believe that the recent advances in machine learning can be useful to better approximate the contributions of sound sources relevant for perceptual attribute prediction such as pleasantness. In this paper, we do so by developing new indicators relying on source recognition models which are built based on deep learning techniques CITE. However, large amounts of annotated data are needed for these architectures to learn to extract relevant information from recordings and to the best of our knowledge such databases do not currently exist in the literature. In order to circumvent this issue, we propose in this paper to study the potential of the use of simulated corpora (CITE) \ml{j'enleverai c'est super ambigu, on a l'impression que l'on annote au hasard :) where sound scenes are composed from semi-randomly generated annotations}.

The contributions of this paper are thus three-fold:
\begin{enumerate}
  \item provide controled acoustic data and corresponding perceptual attributes relevant to the study of urban soundscapes perception
  \item propose numerical means for predicting time of presence of sound sources from raw acoustic data
  \item provide a comparative study of the behavior of the proposed approach with state of the art ones.
\end{enumerate}

First, a perceptual test is conducted where simulated sound scenes are compared to recordings to validate their use in the context of pleasantness prediction. Next, a deep learning architecture is designed and trained that performs multi-label sound source recognition at relevant time scales. Estimations of pleasantness using these predictions are next compared to existing models from physical indicators.

\ml{mettre les noms de section dans ce par}
